### 2019-09-12 initial/notes.tex
writeup WIP

We can investigate the values of $\theta_f$ given some initial mutual
inclination $\hat{s} \cdot \hat{l}_p$ ($J_i = 2\pi\p*{1 - \cos \epsilon}$ where
$\epsilon$ is the initial mutual inclination). To do this, we need to understand
what the possible transitions are. When the separatrix first appears to the
final state, the transitions are $A_2 \to A_3, A_2 \to A_1, A_3 \to A_1, A_3 \to
A_2 \to A_1, A_3 \to A_3$. Let's reproduce the analytical forms of these three
areas below:

%   z0 = eta * np.cos(I)
%   chi = np.sqrt(-np.tan(q4)**3 / np.tan(I) - 1)
%   rho = chi * np.sin(q4)**2 * np.cos(q4) / (
%       chi**2 * np.cos(q4)**2 + 1)
%   T = 2 * chi * np.cos(q4) / (
%       chi**2 * np.cos(q4)**2 - 1)
%   A2 = 8 * rho + 4 * np.arctan(T) - 8 * z0 * np.arctan(1 / chi)
%   A1 = 2 * np.pi * (1 - z0) - A2 / 2
%   A3 = 2 * np.pi * (1 + z0) - A2 / 2
\begin{subequations}\label{se:}
    \begin{align*}
        z_0 &= \eta\cos I &
        \chi &= \sqrt{-\frac{\tan^3\theta_4}{\tan I} - 1},\\
        \rho &= \chi \frac{\sin^2 \theta_4\cos \theta_4}{
            \chi^2 \cos^2\theta_4 + 1} &
        T &= 2\chi \frac{\cos \theta_4}{
            \chi^2 \cos^2\theta_4 - 1},
    \end{align*}
    \begin{align}
        A_2 &= 8\rho + 4\arctan T - 8z_0 \arctan \frac{1}{\chi},\\
        A_1 &= 2\pi\p*{1 - z_0} - \frac{A_2}{2},\\
        A_3 &= 2\pi\p*{1 + z_0} - \frac{A_2}{2}.
    \end{align}
\end{subequations}

>>>>

We can investigate the values of $\theta_f$ given some initial mutual
inclination $\hat{s} \cdot \hat{l}_p \equiv \theta_{sp, i}$ ($J_i = 2\pi\p*{1 -
\cos \epsilon}$ where $\epsilon$ is the initial mutual inclination). To do this,
we need to understand what the possible transitions are. When the separatrix
first appears to the final state, the transitions are $A_2 \to A_3, A_2 \to A_1,
A_3 \to A_1, A_3 \to A_2 \to A_1, A_3 \to A_3$. Let's reproduce the analytical
forms of these three areas below:
\begin{subequations}\label{se:}
    \begin{align*}
        z_0 &= \eta\cos I &
        \chi &= \sqrt{-\frac{\tan^3\theta_4}{\tan I} - 1},\\
        \rho &= \chi \frac{\sin^2 \theta_4\cos \theta_4}{
            \chi^2 \cos^2\theta_4 + 1} &
        T &= 2\chi \frac{\cos \theta_4}{
            \chi^2 \cos^2\theta_4 - 1},
    \end{align*}
    \begin{align}
        A_2 &= 8\rho + 4\arctan T - 8z_0 \arctan \frac{1}{\chi},\\
        A_1 &= 2\pi\p*{1 - z_0} - \frac{A_2}{2},\\
        A_3 &= 2\pi\p*{1 + z_0} - \frac{A_2}{2}.
    \end{align}
\end{subequations}
To determine which of the transition histories we fall under (to get an
analytical prediction for $\cos \theta_f$), we need to determine $\eta_\star$,
so we know how the enclosed $J$ by the trajectory changes due to separatrix
encounter. A guess for numerical root-finding can be put together for
trajectories that encounter from $A_2$ via (inverting the separatrix area
formula for small $\eta$)
\begin{equation*}
    \eta_\star\p*{A_2} = \p*{A_2/16}^2 / \sin I,
\end{equation*}
while for trajectories that encounter from $A_3$ we can solve the quadratic $A_3
= 2\pi \eta \cos I - 8\sqrt{\eta \sin I} = 4\pi - J$ and obtain
\begin{equation*}
    \sqrt{\eta_\star} = \frac{4\sqrt{\sin I} + \sqrt{16\sin I + 2\pi \cos I
        A_3}}{ 2\pi \cos I}.
\end{equation*}
This will often run over $\eta_c$, so we guess the smaller of $\eta_\star,
\eta_c$.

Finally, the $A_3 \to A_2 \to A_1$ pipeline only exists if $ \eta_{A2\max} <
\eta_\star < \eta_c$, the $\eta$ at which $A_2$ is maximized (this maximization
behavior is not captured in my linearized formulas, but can be seen in
\lstinline{99_misc/1_areas.png}). Then, two separatrix crossings occur. Thus, for
the five possible trajectories, we have the following transitions:
\begin{itemize}
    \item $A_2 \to A_3$: This transition happens with probability $\frac{-2\pi
        \eta_{\star} \cos I + 4\sqrt{\eta_{\star}\sin
        I}}{8\sqrt{\eta_{\star}\sin I}}$, so for $0 \leq \eta_{\star} \leq
        \frac{4\sin I}{\pi^2 \cos^2 I}$ this transition is permitted. The
        approximate $\mu_f$ can be estimated by conservation of adiabatic
        invariant
        \begin{align*}
            \eta_\star &\approx \p*{\frac{2\pi\p*{1 - \cos \theta_{sp,i}}}{
                        16}}^2 / \sin I&
                    &\approx \p*{\frac{\pi \theta_{sp, i}^2}{16}}^2/\sin I,\\
            \mu_{f, 23} &\approx \eta_\star \cos I
                    - \frac{4\sqrt{\eta_\star \sin I}}{\pi} &
                &\approx\p*{\frac{\pi \theta_{sp, i}^2}{16}}^2 \cot I
                    - \frac{\theta_{sp, i}^2}{4}.
        \end{align*}

    \item $A_2 \to A_1$: This transition happens with probability $\frac{-2\pi
        \eta_{\star} \cos I + 4\sqrt{\eta_{\star}\sin
        I}}{8\sqrt{\eta_{\star}\sin I}}$ and can happen so long as $\theta_{sp,
        i}$ is small enough to fit inside $A_{crit}$ when the separatrix first
        appears, i.e.\ $2\pi\p*{1 - \cos \theta_{sl, i}} < A_{crit}$ (where
        $\theta_{sl, i}$ is defined in the $\eta \to \infty$ limit, and since
        there is no separatrix the enclosed area is conserved as $\eta$
        approaches $\eta_c$ from above).

        The resulting formulas for $\mu_{f, 21}$ are very similar to those above
        \begin{align*}
            \eta_\star &\approx \p*{\frac{2\pi\p*{1 - \cos \theta_{sp,i}}}{
                        16}}^2 / \sin I&
                    &\approx \p*{\frac{\pi \theta_{sp, i}^2}{16}}^2/\sin I,\\
            \mu_{f, 21} &\approx \eta_\star \cos I
                    + \frac{4\sqrt{\eta_\star \sin I}}{\pi} &
                &\approx\p*{\frac{\pi \theta_{sp, i}^2}{16}}^2 \cot I
                    + \frac{\theta_{sp, i}^2}{4}.
        \end{align*}

    \item $A_3 \to A_3$: This transition occurs if $J_i(\theta_{sl, i}) > 4\pi -
        A_{3, \min}$. This implies the circulation never undergoes a separatrix
        encounter at all, so it remains in $A_3$ forever.

    \item $A_3 \to A_1$: This transition occurs when $\theta_{sl, i}$ satisfies
        $2\pi\p*{1 - \cos \theta_{sl, i}} > A_{crit}$, so the trajectory starts
        in $A_3$. It then has probability to hop to $A_1$ (TODO).

    \item $A_3 \to A_2 \to A_1$: This transition also occurs when $J_i$ is
        $> A_{crit}$, but has one further requirement, that $\eta_{\star} >
        \argmax A_2(\eta)$, such that $A_2$ is expanding during the first
        separatrix encounter. The bound on this is not exactly analytic though.
\end{itemize}
